---
title: "p8105_hw6"
author: "Renee Wang UNI:hw2851"
date: "12/4/2021"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)


```
# Problem 1 

read and clean data
```{r, include=FALSE}
birth_weight_df = 
  read_csv("./birthweight.csv") %>%
  na.omit() %>%
  mutate(
   babysex = recode_factor(babysex, "1" = "male", "2" = "female"),
    frace = recode_factor(frace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other", "9" = "Unkown"),
    mrace = recode_factor(mrace, "1" = "White", "2" = "Black", "3" = "Asian", "4" = "Puerto Rican", "8" = "Other"),
    malform = recode_factor(malform, "0" = "absent", "1" = "present")
  )

```
```{r}
birth_weight_df
```

After cleaning the data , we have seen variable "pnumsga" and "pnumlbw" are all 0 across the dataset, and variable "parity"and "malform" have less than five values included, so we exclude those variables and take the others as the predictors.

```{r, include=FALSE}
birth_weight_df %>%
  select()
  
```

Low birth weight is a major public health problem. As such, there is a large number of research literature studying the underlying factors that may affect birth weight, so we decide to choose the predictors based on existing research.
In particular, we referred to this article: Metgud, Chandra S., Vijaya A. Naik, and Maheshwar D. Mallapur. "Factors affecting birth weight of a newborn–a community based study in rural Karnataka, India." PloS one 7.7 (2012): e40040.
The authors identified a few factors that are significantly associated with birth weight: exposure to passive smoking, age at first pregnancy, birth interval, weight gain during pregnancy, etc. We checked this list of factors against the given dataset and decide to use the following variables as predictors: babysex, delwt, fincome, malform, momage, smoken and wtgain.

The next step is to decide a model to use. Based on the literature, many of these factors has non-linear impact on birth weight. Thus, we decide to use glm instead of lm.

```{r}
fit = glm(bwt ~ babysex + delwt + fincome + malform + momage + smoken + wtgain, data = birth_weight_df)
birth_weight_df %>%
  modelr::add_residuals(fit) %>%
  modelr::add_predictions(fit) %>%
  ggplot(aes(x = pred, y = resid)) + geom_point() + geom_smooth(method='lm') +
  labs(
    x = "Fitted values",
    y = "Residuals",
    title = "Birth weight prediction model residuals against fitted values"
  )
```
Looking at the plot, residuals are centered around 0 and are mostly symmetrical. This can also be seen from the lm regression line we added to the plot. There are a few outliers, but for most of the data points we are making a reasonable prediction with a small residual.

Now we compare our model against the other two models, using cross validation.
```{r}
cv_df =
  crossv_mc(birth_weight_df, 100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>%
  mutate(
    our_method = map(train, ~glm(bwt ~ babysex + delwt + fincome + malform + momage + smoken + wtgain, data = .x) ),
    blength_gaweeks_method = map(train, ~glm(bwt ~ blength + gaweeks, data = .x)),
    bhead_blength_babysex_method = map(train, ~glm(bwt ~ bhead * blength * babysex, data = .x))
  ) %>%
  mutate(
    rmse_our = map2_dbl(our_method, test, ~rmse(model = .x, data = .y)),
    rmse_blength_gaweeks = map2_dbl(blength_gaweeks_method, test, ~rmse(model = .x, data = .y)),
    rmse_bhead_blength_babysex = map2_dbl(bhead_blength_babysex_method, test, ~rmse(model = .x, data = .y))
  )
```

We post-process the cross-validation result a bit to get the violin plot of the RMSE errors of the three models:
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>%
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```
From the plot, it's clear that the model using head circumference, length, sex, and all interactions is the best-performing one, with mean RMSE of `r mean(cv_df$rmse_bhead_blength_babysex)`; the model using length at birth and gestational age is the second best performing, with mean RMSE of `r mean(cv_df$rmse_blength_gaweeks)`. Our model, unfortunately, performs worse than the other two models with mean RMSE of `r mean(cv_df$rmse_our)`.